Global patterns of forest autotrophic carbon fluxes

### 2020-08-06 manuscript/manuscript-draft.tex
finished first pass at revising discussion

Forest C fluxes decrease with latitude
\citep[e.g.,][]{luyssaert_co_2007, piao_forest_2010, gillman_latitude_2015, li_mapping_2019},
but it remains unclear whether and how the shape of this relationship
varies among fluxes. For instance, gross primary productivity (\(GPP\))
increases continuously with decreasing latitude and is indisputably
highest in tropical forests
\citep{luyssaert_co_2007, beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
In contrast, some studies have suggested that net primary productivity
(\(NPP\)), or its aboveground portion (\(ANPP\)), exhibits a less
distinct increase from temperate to tropical forests
\citep{luyssaert_co_2007}--or even a decrease \citep[but see
\citet{gillman_latitude_2015}]{huston_global_2009}. A shallower increase
in \(NPP\) than in \(GPP\) with decreasing latitude would align with the
suggestion that tropical forests tend to have low carbon use efficiency
\citep[\(CUE\)=
\(NPP\)/\(GPP\);][]{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
However, accurately comparing relationships of multiple forest C fluxes
across latitude requires a large database with standardization for
methodology, stand history, and management, which to date has not been
possible.

@@

\hspace{1em}H2.3. Temperature and precipitation interactively shape C fluxes.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\

@@

Our analysis of a large global database (ForC) reveals how autotrophic
carbon fluxes in mature forests vary with latitude and climate on a
global scale. We show that, across all nine forest autotrophic C fluxes
analyzed, C cycling decreases continually with latitude (\emph{H1.1};
Fig. 2)--a finding that confirms multiple previous studies but
contradicts the idea that productivity of temperate forests rivals or
even exceeds that of tropical forests
\citep{luyssaert_co_2007, huston_global_2009}. At this global scale, C
fluxes increase approximately in proportion to one another, with
component fluxes summing appropriately to larger fluxes and no
detectable differences in allocation across latitude or climates (Figs.
2, 4, S3). However, climate explains lower proportions of variability
among some of the subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\),
\(BNPP\) \(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). Latitudinal
variation in C fluxes is primarily attributable to temperature-related
variables (\emph{H3, H4}), particularly MAT and temperature seasonality
(Figs. 3-4). Water availability is also influential, but generally of
secondary importance across the climate space occupied by forests (Figs.
3-4). While growing season length explained signficant variation in all
C fluxes (\emph{H4}; Figs. 4f, S6-S7), it did not improve upon MAT or
temperature seasonality as a predictor. Within the growing season, the
influence of climate on C cycling is smaller but still significant for a
number of C fluxes (\emph{H5}; Fig. S9; Table S4).

These findings clarify how forest C fluxes vary with latitude and
climate on a global scale. Past studies have differed in their
conclusions regarding the relationship between C fluxes and latitude or
its correlates (Table 1, \emph{H1})--quite possibly because of lack of
standardization with respect to stand age and disturbance history. Our
findings indicate that, among mature, undisturbed stands, forest C
fluxes are unambiguously highest in the tropical regions, and the
relationship with latitude is approximately linear (Fig. 2). This
contrasts with the suggestion that productivity of temperate forests is
similar to that of tropical forests \citep{huston_global_2009}. Compared
to tropical forests, the temperate forest biome has experienced more
widespread anthropogenic disturbance and has a larger fraction of
secondary stands \citep{potapov_mapping_2008, poulter_global_2018}, so
analyses comparing across latitudinal gradients without controlling for
stand age risk confounding age with biome effects. In addition, because
carbon allocation varies with stand age
\citep{de_lucia_forest_2007, anderson-teixeira_altered_2013, doughty_what_2018},
age differences may introduce systematic biases into analyses of C
fluxes across latitude or global climatic gradients. For example, woody
productivity tends to be higher in rapidly aggrading secondary stands
than in old-growth forests, where proportionally more C is allocated to
respiration
\citep{de_lucia_forest_2007, piao_forest_2010, doughty_what_2018, kunert_understanding_2019}.

We show that C fluxes are broadly consistent in their responses to
climate drivers on the global scale, with at most modest trends in C
allocation among the variable pairs tested (Figs. 2, S3). This parallels
the observation that C allocation across multiple C fluxes varies little
with respect to climate along a steep tropical elevational gradient
\citetext{\citealp{malhi_variation_2017}; \citealp[but
see][]{moser_elevation_2011}}, and is not surprising given that trees
face universal constraints in carbon allocation (\textbf{REFS}). The one
trend in allocation that emerged from our analysis was a tendency for
greater C allocation aboveground--and to foliage in particular--in warm
tropical climates (Fig. S3). This is consistent with observations of
increasing allocation to roots with declining temperature across a
tropical elevational gradient \citep{moser_elevation_2011}, and with
observations and theory predicting relatively higher belowground carbon
allocation at higher latitudes \citep{gill_belowground_2016}.
\textbf{Krista -- I took out the ANPP:NPP graphs from the SI after
looking back at what Helene said in issue \#77 -- let me know if you
think it's worth putting them back in wrt this point} It is also
consistent with findings that as forest productivity increases,
proportionally more carbon is allocated to \(ANPP_{foliage}\) relative
to \(BNPP_{fine.root}\) \citep{chen_allometry_2019} or \(ANPP_{stem}\)
\citep{hofhansl_new_2015}. \textbf{( check / comment on
\citet{litton_carbon_2007} )}

One previously hypothesized trend that was not supported by our analysis
was that tropical forests tend to have low carbon use efficiency
(\(CUE\)= \(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which is based
on observations of low \(CUE\) in old-growth tropical forests relative
to (mostly younger) extratropical forests
\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
Our analysis, limited to mature forests, shows no such trend (Fig. S3).
\(CUE\) is known to decline with forest age
\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}, but
appears to be roughly independent of \(GPP\) \citep{litton_carbon_2007}.
(\emph{This finding may have some important implications for modeling
and our broader understanding. I'm not sure offhand how much theory has
been built on the idea of low CUE in tropical forests because of warm
temperatures\ldots Helene is not aware of any.})

\textbf{update this paragraph
(\href{https://github.com/forc-db/Global_Productivity/issues/78}{issue
78})} One interesting observation was that climate tends to explain more
variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\)) than in
subsidiary fluxes (\(BNPP_{fine.root}\), \(R_{root}\), \(ANPP_{stem}\))
as quantified by R\textsuperscript{2} (Fig. 2; Table S2). There are two,
non-exclusive, potential explanations for this. First, it may be that
methodological variation is larger relative to flux magnitude for some
of the subsidiary fluxes. Belowground fluxes in particular are difficult
to quantify, and measurement methods for the belowground fluxes
considered here may use fundamentally different approaches in different
sites (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential coring
for \(BNPP_{fine.root}\); root exclusion, stable isotope tracking, or
gas exchange of excised roots for \(R_{root}\)), and sampling depth is
variable and often insufficient to capture the full soil profile.
\(ANPP_{stem}\), which is also poorly explained by latitude or climate,
is more straightforward to measure but is subject to variability
introduced by differences such as biomass allometries applied and
minimum plant size sampled \citep{clark_measuring_2001}. However,
methodological variation and uncertainty affect all of fluxes considered
here, and some of the larger fluxes that vary more strongly with respect
to climate (\(ANPP\), \(NPP\)) are estimated by summing uncertain
component fluxes. Second, differences among variables in the proportion
of variation explained by climate may be attributable to more direct
climatic control over \(GPP\) than subsidiary fluxes. That is,
subsidiary fluxes may be shaped by climate both indirectly through its
influence on \(GPP\) and respiration and directly through any climatic
influence on C allocation, as well as many other local- and
regional-scale factors (\textbf{REFS}).

The latitudinal gradient in forest C flux (Fig. 2) is driven primarily
by temperature-related climate variables, the effects of which are
moderated by moisture availability (Table 1, \emph{H2-H3}; Figs. 3-4).
MAT and MAP have long been identified as primary global-scale drivers of
C fluxes
\citetext{\citealp{lieth_primary_1973}; \citealp[\textbf{REFS};][]{taylor_temperature_2017}}.
It is not appropriate to attempt to identify individual mean annual
climate variables as mechanistic drivers of C fluxes because many
climate variables co-vary across the latitudinal gradient (Fig. S2),
because climatic drivers affect forest carbon flux on much shorter time
scales than can be captured by annual climate summary variables, and
because both climatic conditions and C flux vary intra- and
inter-annually around the long-term means. However, it remains
informative to consider these relationships. Among the
temperature-related climate variables, \(MAT\) is generally the most
strongly correlated with C fluxes (Table S2)--perhaps in part because
site-specific MAT is recorded for the majority of sites in ForC, whereas
other variables were extracted from global gridded data products,
introducing spatial scaling errors.

This finding supports the continued focus on MAT as a
primary--\emph{albiet} not mechanistic--correlate of C fluxes. The
effects of \(MAT\) are modified by moisture availability, with reduced C
fluxes under relatively dry conditions (\emph{i.e.}, low precipitation;
high vapour pressure deficit) and sometimes under very high
precipitation (Figs. 3-4). \textbf{BBL: I feel like this is pretty
repetitive of previous paragraph.} The observed positive interaction
between MAT and MAP for \(ANPP_{stem}\) on the global scale (Fig. 3) is
consistent with an analysis showing a similar interaction for \(ANPP\)
in tropical forests, also with a cross-over point at
\textasciitilde20\(\circ\)C \citep{taylor_temperature_2017}.
\emph{(\citet{sullivan_sensitivity_2020} show a decline in
\(ANPP_{stem}\) in sites with high max temperatures and limited
moisture.)} However, we detect no such interaction for \(ANPP\) or most
other C fluxes, and we find a contrasting negative interaction for
\(NPP\) (Fig. 3), suggesting that more data are required to sort out
potential differences in the interactive effects of MAT and MAP on C
fluxes in the tropics.

Forest autotrophic C fluxes decline with temperature seasonality (Table
1, \emph{H4}; Fig. 4e), and are minimal during cold- or dry- dormant
seasons. To account for this, a number of analyses seeking to
characterize global-scale effects of climate on productivity have
examined the relationship of C flux per month of the growing season with
growing season climatic conditions \citep[Table 1,
\emph{H5};][]{kerkhoff_plant_2005, anderson_temperature-dependence_2006, enquist_adaptive_2007, michaletz_convergence_2014}.
The sort of simple metric that has been used to define growing season at
a global scale \citep{kerkhoff_plant_2005} is coarse with respect to
temperature because it's calculated on a monthly timescale and
problematic with respect to moisture because it doesn't capture temporal
lags between precipitation and plant water use caused by storage in soil
or snow. We found that a temperature-defined growing season length had
strong positive correlation with C fluxes (Fig. 4f), but was never the
best. Dividing annual fluxes by growing season length to yield average
flux per growing season month removed the majority of climate-related
variation, supporting the idea that the latitudinal gradient in carbon
flux is attributable more to shorter growing seasons at high latitudes
than to inherently lower rates of photosynthesis or respiration by
high-latitude forests \citep{enquist_adaptive_2007}. However, there
remained a number of significant correlations with growing season
climatic conditions, indicating that climatic conditions remain
influential within the growing season. We conclude that while correcting
for growing season length takes analyses a step closer to mechanistic
linkage of instantaneous C flux rates to environmental conditions, it
remains crude relative to the timescales on which climate affects plant
metabolism, and does not advance statistical predictive power.
Mechanistic accounting for climatic effects on global forest carbon flux
patterns instead requires models representing physiologically meaningful
timescales \citep[e.g., \textbf{REFS};][]{longo_biophysics_2019}.

Our analysis clarifies how forest autotrophic carbon fluxes vary with
latitude and climate on a global scale, with some important implications
for how forest carbon cycling relates to climate and, by extension, how
it is likely to respond to climatic warming. Our findings show that
higher temperatures with similar moisture availability result in a
generalized acceleration of forest C cycling (Figs. 2-3). This is
consistent with observations of continental- to global-scale increases
over time in \(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
\citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
C cycle components not considered here--tree mortality
\citep{brienen_long-term_2015, mcdowell_drivers_2018}, soil respiration
\citep{bond-lamberty_global_2010}, and heterotrophic soil respiration
\citep{bond-lamberty_globally_2018}. However, increasing C flux rates
are by no means universal
\citep[e.g.,][]{rutishauser_testing_2020, hubau_asynchronous_2020},
likely because other factors are at play, including changes to other
aspects of climate, atmospheric pollution (CO\textsubscript{2},
SO\textsubscript{2}, NO\textsubscript{x}), and local disturbances.\\
In addition, some tropical forests are \textbf{(discuss new Sullivan
paper that finds higher max temperatures associated to lower
ANPP\_stem)} \citep{sullivan_sensitivity_2020} Moreover, forest
ecosystem responses to climatic changes outside the temperature range to
which forest communities are adapted and acclimatized will not
necessarily parallel responses across geographic gradients in climate.
Nevertheless, understanding of the fundamental climatic controls on C
cycling sets a firmer foundation for understanding forest C cycle
responses to accellerating climate change.

>>>>

Forest C fluxes decrease with latitude
\citep[e.g.,][]{luyssaert_co_2007, piao_forest_2010, gillman_latitude_2015, li_mapping_2019},
but studies have differed in their conclusions regarding the shape of
this relationship - quite possibly because of lack of standardization
with respect to methodology and stand history. For instance, studies
agree that gross primary productivity (\(GPP\)) increases continuously
with decreasing latitude and is indisputably highest in tropical forests
\citep{luyssaert_co_2007, beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
In contrast, some studies have suggested that net primary productivity
(\(NPP\)), or its aboveground portion (\(ANPP\)), exhibits a less
distinct increase from temperate to tropical forests
\citep{luyssaert_co_2007}--or even a decrease \citep[but see
\citet{gillman_latitude_2015}]{huston_global_2009}. A shallower increase
in \(NPP\) than in \(GPP\) with decreasing latitude would align with the
suggestion that tropical forests tend to have low carbon use efficiency
\citep[\(CUE\)=
\(NPP\)/\(GPP\);][]{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
Such differences among C fluxes their relationship to latitude could
have profound implications for our understanding of the C cycle and its
climate sensitivity. However, until recently the potential to compare
latitudinal trends across C fluxes has been limited by lack of a large
database with standardization for methodology, stand history, and
management \citep{anderson-teixeira_forc_2018}.

@@

\hspace{1em}H2.3. Temperature and precipitation jointly shape C fluxes.\textsuperscript{5} & (yes) & yes & yes & yes & yes & - & - & - & yes & - & Fig. 3, Table S3\\

@@

Our analysis of a large global database (ForC) clarifies how autotrophic
C fluxes in mature forests vary with latitude and climate on a global
scale. We show that, across all nine variables analyzed, annual C flux
decreases continually with latitude (Fig. 2)--a finding that confirms
multiple previous studies but contradicts the idea that productivity of
temperate forests rivals or even exceeds that of tropical forests
\citep{luyssaert_co_2007, huston_global_2009}. At this global scale, C
fluxes increase approximately in proportion to one another, with
component fluxes summing appropriately to larger fluxes and no
detectable differences in allocation across latitude or climates (Figs.
2, 4, S3). Similarly, we show broad - \emph{albeit} not complete -
consistency of climate responses across C fluxes, with the observed
latitudinal variation primarily attributable to temperature and its
seasonality (Figs. 3-4). Water availability is also influential, but of
secondary importance across the climate space occupied by forests (Figs.
3-4). Contrary to prior suggestions that the majority of variation in C
cycling is driven by the length of the growing season (\textbf{REFS}),
we find modest explanatory power of growing season length and small but
sometimes significant influence of climate within the growing season
(Figs. 4f,S6-S9). Together, these findings yield a unified understanding
of climate's influence on forest C cycling.

Our findings indicate that, among mature, undisturbed stands, forest C
fluxes are unambiguously highest in the tropical regions, and the
relationship with both latitude and \(MAT\) is approximately linear
(Figs. 2, 4). This contrasts with the suggestion that C fluxes (e.g.,
\(NPP\), \(ANPP\), \(ANPP_{stem}\)) of temperate forests are similar to
or even greater than that of tropical forests
\citep{luyssaert_co_2007, huston_global_2009}. Previous indications of
such a pattern may have been an artifact of differences in stand age
across biomes. Compared to tropical forests, the temperate forest biome
has experienced more widespread anthropogenic disturbance and has a
larger fraction of secondary stands
\citep{potapov_mapping_2008, poulter_global_2018}, so analyses comparing
across latitudinal gradients without controlling for stand age risk
confounding age with biome effects. Because carbon allocation varies
with stand age
\citep{de_lucia_forest_2007, anderson-teixeira_altered_2013, doughty_what_2018},
age differences may introduce systematic biases into analyses of C
fluxes across latitude or global climatic gradients. For example, woody
productivity tends to be higher in rapidly aggrading secondary stands
than in old-growth forests, where proportionally more C is allocated to
respiration and non-woody productivity
\citep{de_lucia_forest_2007, piao_forest_2010, doughty_what_2018, kunert_understanding_2019}.
Thus, findings that temperate forest productivity rivals that of
tropical forests are likely an artifact of different forest ages across
biomes.

We show that C fluxes are broadly consistent in their responses to
climate drivers on the global scale, with no trends in C allocation
among the variable pairs tested (Figs. 2, 4, S3). This parallels the
observation that C allocation across multiple C fluxes varies little
with respect to climate along a steep tropical elevational gradient
\citetext{\citealp{malhi_variation_2017}; \citealp[but
see][]{moser_elevation_2011}}, and is not surprising given that trees
face universal constraints in carbon allocation (\textbf{REFS}). We find
no trend in the allocation of \(GPP\) between production and respiration
across latitude or climate (\(NPP\):\(R_{auto}\); Fig. S3), refuting the
idea that tropical forests have anomalously low \(CUE\)
\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
Rather, differences in \(CUE\) between old-growth tropical forests
relative to (mostly younger) extratropical forests are likely an
artifact of comparing stands of different age, as \(CUE\) is known to
decline with forest age
\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}.
Another previously observed pattern for which we find no support is a
tendency for belowground C allocation to decrease with increasing
temperature \citep{moser_elevation_2011, gill_belowground_2016}; rather,
we observe no trends in allocation between \(ANPP\) and \(BNPP\) across
latitudes. Failure to detect signficant tends in C allocation with
respect to climate in this analysis does not imply that none exist;
rather, it suggests that

Despite the broad consistency of climate responses across C fluxes,
climate explains lower proportions of variability among some of the
subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\), \(BNPP\)
\(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). There are two,
non-exclusive, potential explanations for this. First, it may be that
methodological variation is larger relative to flux magnitude for some
of the subsidiary fluxes. Belowground fluxes in particular are difficult
to quantify, and measurement methods for the belowground fluxes
considered here may use fundamentally different approaches in different
sites (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential coring
for \(BNPP_{fine.root}\); root exclusion, stable isotope tracking, or
gas exchange of excised roots for \(R_{root}\)), and sampling depth is
variable and often insufficient to capture the full soil profile.
\(ANPP_{stem}\), which is also poorly explained by latitude or climate,
is more straightforward to measure but is subject to variability
introduced by differences such as biomass allometries applied and
minimum plant size sampled \citep{clark_measuring_2001}. However,
methodological variation and uncertainty affect all of fluxes considered
here, and some of the larger fluxes that vary more strongly with respect
to climate (\(ANPP\), \(NPP\)) are estimated by summing uncertain
component fluxes. Second, differences among variables in the proportion
of variation explained by climate may be attributable to more direct
climatic control over \(GPP\) than subsidiary fluxes. That is,
subsidiary fluxes may be shaped by climate both indirectly through its
influence on \(GPP\) and respiration and directly through any climatic
influence on C allocation, as well as many other local- and
regional-scale factors (\textbf{REFS}).

Temperature and its seasonality were the primary drivers of C flux on
the global scale (Figs. 2-4), consistent with a long legacy of research
identifying temperature as a primary driver of forest ecosystem C
cycling {[}e.g., \citet{lieth_primary_1973}; \textbf{REFS}{]}. We find
little evidence of any non-linearity in temperature's influence on C
flux. The relationship of all fluxes to \(MAT\) as an individual driver
were best described by a linear function (Table S2) - with the exception
of \(BNPP\), whose response to \(MAT\) was close to linear (Fig. 4a).
This result contrasts with both the idea that C flux responds
exponentially to temperature (\textbf{REFS}) and with the idea that
fluxes saturate with \(MAT\) below approximately 25\(^\circ\)C MAT
(\textbf{REFS}). It remains possible that flux declines above this
threshold
\citep{larjavaara_temperature_2012, sullivan_sensitivity_2020}, as is
also consistent with tree-ring records indicating that tropical tree
growth declines at high temperatures
\citep[e.g.,][]{vlam_temperature_2014}. However, these higher
temperatures also tend to be associated with high PET and VPD, both of
which are associated with reduced C fluxes (Figs. 4c-d, S4-S5). Indeed,
while temperature responses dominate at this global scale and with the
climate space occupied by forests, the effects of temperature are
moderated by moisture availability (Table 1; Figs 3-4). Specifically, C
fluxes are reduced under relatively dry conditions (\emph{i.e.}, low
\(MAP\); high \(VPD\)) and sometimes under very high precipitation
(Figs. 3-4). The observed positive interaction between \(MAT\) and
\(MAP\) for \(ANPP_{stem}\) on the global scale (Fig. 3) is consistent
with an analysis showing a similar interaction for \(ANPP\) in tropical
forests, also with a cross-over point at \textasciitilde20\(\circ\)C
\citep{taylor_temperature_2017}.\\
However, we detect no such interaction for \(ANPP\) or most other C
fluxes, and we find a contrasting negative interaction for \(NPP\) (Fig.
3), suggesting that more data are required to sort out potential
differences in the interactive effects of MAT and MAP on C fluxes in the
tropics.

Forest autotrophic C fluxes decline with temperature seasonality (Table
1, \emph{H4}; Fig. 4e), and are minimal during cold- or dry- dormant
seasons. To account for this, a number of analyses seeking to
characterize global-scale effects of climate on productivity have
examined the relationship of C flux per month of the growing season with
growing season climatic conditions \citep[Table 1,
\emph{H5};][]{kerkhoff_plant_2005, anderson_temperature-dependence_2006, enquist_adaptive_2007, michaletz_convergence_2014}.
The sort of simple metric that has been used to define growing season at
a global scale \citep{kerkhoff_plant_2005} is coarse with respect to
temperature because it is calculated on a monthly timescale and
problematic with respect to moisture because it doesn't capture temporal
lags between precipitation and plant water use caused by storage in soil
or snow. We found that a temperature-defined growing season length had
strong positive correlation with C fluxes (Fig. 4f), but was never the
best univariate predictor. Dividing annual fluxes by growing season
length to yield average flux per growing season month removed the
majority of climate-related variation, supporting the idea that the
latitudinal gradient in carbon flux is attributable more to shorter
growing seasons at high latitudes than to inherently lower rates of
photosynthesis or respiration by high-latitude forests
\citep{enquist_adaptive_2007}. However, there remained a number of
significant correlations with growing season climatic conditions,
indicating that climatic conditions remain influential within the
growing season (Figs. S8-S9). We conclude that while correcting for
growing season length takes analyses a step closer to mechanistic
linkage of instantaneous C flux rates to environmental conditions, it
remains crude relative to the timescales on which climate affects plant
metabolism, and does not advance statistical predictive power.
Mechanistic accounting for climatic effects on global forest carbon flux
patterns instead requires models representing physiologically meaningful
timescales
\citep[e.g.,][]{medvigy_mechanistic_2009, longo_biophysics_2019}.

Our analysis clarifies how forest autotrophic carbon fluxes vary with
latitude and climate on a global scale, with some important implications
for how forest carbon cycling relates to climate and, by extension, how
it is likely to respond to climatic warming. To the extend that patterns
across broad scale climatic gradients can foretell how ecosystem
responses to climate change, our findings suggest that higher
temperatures with similar moisture availability result in a generalized
acceleration of forest C cycling (Figs. 2-3). This is consistent with
observations of continental- to global-scale increases over time in
\(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
\citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
C cycle components not considered here--tree mortality
\citep{brienen_long-term_2015, mcdowell_drivers_2018}, soil respiration
\citep{bond-lamberty_global_2010}, and heterotrophic soil respiration
\citep{bond-lamberty_globally_2018}. However, increasing C flux rates
are by no means universal
\citep[e.g.,][]{rutishauser_testing_2020, hubau_asynchronous_2020},
likely because other factors are at play, including changes to other
aspects of climate, atmospheric pollution (CO\textsubscript{2},
SO\textsubscript{2}, NO\textsubscript{x}), and local disturbances.\\
Moreover, forest ecosystem responses to climatic changes outside the
temperature range to which forest communities are adapted and
acclimatized will not necessarily parallel responses across geographic
gradients in climate. Indeed, tree-ring studies from forests around the
world indicate that tree growth rates - along with \(ANPP_{stem}\) and
possibly other ecosystem C fluxes - respond negatively to temperature
\citep{sniderhan_growth_2016, helcoski_growing_2019}. Furthermore, in
the tropics, climate change will push forests beyond any contemporary
climate, and there are some indications that this could reduce C flux
rates \citep{sullivan_sensitivity_2020} (\textbf{REFS}). Thus, while
further research is required to understand the extent to which forest
responses to climate change will track global gradients, and the time
scale on which they will do so, understanding the fundamental climatic
controls on annual C cycling in Earth's forests sets a firmer foundation
for understanding forest C cycle responses to accelerating climate
change.
